---
title: "Random Forest Models"
author: "AE"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bremen Big Data Challenge
## Random Forest Models
This file contains only the random forest models shown in both the Main PresentationCode file and the Matching Dates file. 

### Recap
The challenge given to contestants was to take wind farm data collected from January, 2016 to June 2017, and the predict the wind farm's output in July - December of 2017.

The data includes:
* Wind speed and direction at different elevations (All measurements were collected at various heights. All measurements were only taken at certain intervals, and as such certain observations have been linearly interpolated.) 
* A boolean variable which reflects whether the given observation was interpolated or not 
* Capacity of the farm at a given time (This capacity variable is very nearly constant, and most likely only shows variation due to maitainence or downages at the farm.)
* Output of the farm (The original "challenge" data set presented to the contestants did not have any values in the output variable. Outside of the voiding of the output column, the challenge set and validation set are identical.)

### Libraries
```{r Libs}
library(foreign)
library(MASS)
library(car)
library(corrplot)
library(rpart) #for trees
library(ipred) #for bagging
library(randomForest) #for random forest
library(readr) #read in data
library(gbm) #for gradient boosting
library(rms)
library(gam)
library(ISLR)
library(caTools)
library(circular)
library(openair)
```

### Data Import and Preprocessing

```{r Data, results = 'hide'}
#Training Data
train <- read_csv("Data/train.csv")
View(train)
summary(train)
sum(is.na(train)) #No missings

#Reviewing the date variable
class(train$Datum)
#Altering the location, so info is saved in English
Sys.setlocale("LC_ALL", "en_US.UTF-8")
#converting to date format
train$Datum <- as.Date(train$Datum, format = c("%Y-%m-%d"))

#Saving a copy of the training set without date
Train.nodate <- train[,-1]

#Validation Data
eval <- read_csv("Data/eval.csv")
View(eval)
summary(eval)
sum(is.na(eval)) #no missings

#Converting the validation set to a dataframe
class(eval)
eval <- as.data.frame(eval)

#converting date variable to type date
class(eval$Datum)
eval$Datum <- as.Date(eval$Datum, format = c("%Y-%m-%d"))

#Results Data
load("Data_out/results.Rda")
```

### Data Manipulation, Feature Creation, and Spliting the Training Data

```{r Features}
#Adding Variables for Analysis
#Adding month variable
train$month <- months(train$Datum)
train$season <- 0
train$season[train$month == "May"] = 1
train$season[train$month == "June"] = 1
train$season[train$month == "July"] = 1
train$season[train$month == "August"] = 1
train$season[train$month == "September"] = 1
train$season[train$month == "October"] = 1

train$month[train$month == "January"] = 1
train$month[train$month == "February"] = 2
train$month[train$month == "March"] = 3
train$month[train$month == "April"] = 4
train$month[train$month == "May"] = 5
train$month[train$month == "June"] = 6
train$month[train$month == "July"] = 7
train$month[train$month == "August"] = 8
train$month[train$month == "September"] = 9
train$month[train$month == "October"] = 10
train$month[train$month == "November"] = 11
train$month[train$month == "December"] = 12
train$month <- as.numeric(train$month)

eval$month <- months(eval[,1])
eval$season <- 0
eval$season[eval$month == "May"] = 1
eval$season[eval$month == "June"] = 1
eval$season[eval$month == "July"] = 1
eval$season[eval$month == "August"] = 1
eval$season[eval$month == "September"] = 1
eval$season[eval$month == "October"] = 1

eval$month[eval$month == "January"] = 1
eval$month[eval$month == "February"] = 2
eval$month[eval$month == "March"] = 3
eval$month[eval$month == "April"] = 4
eval$month[eval$month == "May"] = 5
eval$month[eval$month == "June"] = 6
eval$month[eval$month == "July"] = 7
eval$month[eval$month == "August"] = 8
eval$month[eval$month == "September"] = 9
eval$month[eval$month == "October"] = 10
eval$month[eval$month == "November"] = 11
eval$month[eval$month == "December"] = 12
eval$month <- as.numeric(eval$month)

#Descriptives for the wind direction and speed
train$winddirectionmean <- ((train$Windrichtung48M+train$Windrichtung100M+train$Windrichtung152M)/3)
train$winddirectionvar <- (((train$Windrichtung48M-train$winddirectionmean)^2+(train$Windrichtung100M-train$winddirectionmean)^2+(train$Windrichtung152M-train$winddirectionmean)^2)/3)
train$windspeedmean <- ((train$Windgeschwindigkeit48M+train$Windgeschwindigkeit100M+train$Windgeschwindigkeit152M)/3)
train$windspeedvar <- (((train$Windgeschwindigkeit48M-train$windspeedmean)^2+(train$Windgeschwindigkeit100M-train$windspeedmean)^2+(train$Windgeschwindigkeit152M-train$windspeedmean)^2)/3)

eval$winddirectionmean <- ((eval$Windrichtung48M+eval$Windrichtung100M+eval$Windrichtung152M)/3)
eval$winddirectionvar <- (((eval$Windrichtung48M-eval$winddirectionmean)^2+(eval$Windrichtung100M-eval$winddirectionmean)^2+(eval$Windrichtung152M-eval$winddirectionmean)^2)/3)
eval$windspeedmean <- ((eval$Windgeschwindigkeit48M+eval$Windgeschwindigkeit100M+eval$Windgeschwindigkeit152M)/3)
eval$windspeedvar <- (((eval$Windgeschwindigkeit48M-eval$windspeedmean)^2+(eval$Windgeschwindigkeit100M-eval$windspeedmean)^2+(eval$Windgeschwindigkeit152M-eval$windspeedmean)^2)/3)

#weighted average direction
train$winddirectionweightedmean <- (((train$Windrichtung48M*train$Windgeschwindigkeit48M)+(train$Windrichtung100M*train$Windgeschwindigkeit100M)+(train$Windrichtung152M*train$Windgeschwindigkeit152M))/(train$Windgeschwindigkeit48M+train$Windgeschwindigkeit100M+train$Windgeschwindigkeit152M))

eval$winddirectionweightedmean <- (((eval$Windrichtung48M*eval$Windgeschwindigkeit48M)+(eval$Windrichtung100M*eval$Windgeschwindigkeit100M)+(eval$Windrichtung152M*eval$Windgeschwindigkeit152M))/(eval$Windgeschwindigkeit48M+eval$Windgeschwindigkeit100M+eval$Windgeschwindigkeit152M))

eval <- as.data.frame(eval)
train <- as.data.frame(train)
```

### Random Forest Models

```{r RF}
#Full formula model
rforest <- randomForest(Output~., data=train)
plot(rforest)
rforest$rsq
importance(rforest, type=1)
importance(rforest, type=2)

#Best formula Model
rforest.1 <- randomForest(Output~ Datum+Windgeschwindigkeit48M+Windgeschwindigkeit100M+Windgeschwindigkeit152M+Windrichtung48M+Windrichtung100M+Windrichtung152M
                          +Windgeschwindigkeit100MP40+Windgeschwindigkeit100MP50+Windgeschwindigkeit100MP70+Windgeschwindigkeit100MP90+Verfügbare_Kapazität, data=train)
plot(rforest.1)
rforest.1$rsq
```

#### Predictions
```{r Preds_rf}
#Full model forest
print("Full Random Forest Model")
predict.rforest <- round(predict(rforest, newdata=eval), digits=0)
(sum((eval$Output-predict.rforest)^2)/nrow(eval))^(1/2)
error_rf <- (sum(abs(predict.rforest-eval$Output)))/sum(eval$Verfügbare_Kapazität)
error_rf

#best model
print("Best Model Formula")
predict.rforest1 <- round(predict(rforest.1, newdata=eval), digits=0)
(sum((eval$Output-predict.rforest1)^2)/nrow(eval))^(1/2)
error_rfbest <- (sum(abs(predict.rforest1-eval$Output)))/sum(eval$Verfügbare_Kapazität)
error_rfbest

results <- 
  rbind(results, 
        list("Full Formula Random Forest",
             error_rf),
        list("Best Formula Random Forest",
             error_rfbest))
```

#### Tuneing the Random Forest Models
```{r RF_tune}
tuned <- tuneRF(train[,-19], train[,19], stepFactor = 2, improve=0.05, plot=TRUE)
```

It appears that the best mtry value, the number of variable tried at each split of the tree, is 8 variables.

#### New Tuned Model
```{r Tuned_RF}
rforest.2 <- 
  randomForest(Output~ Datum + 
                 Windgeschwindigkeit48M +
                 Windgeschwindigkeit100M +
                 Windgeschwindigkeit152M + 
                 Windrichtung100M +
                 Windrichtung152M +
                 Windgeschwindigkeit100MP10+
                 Windgeschwindigkeit100MP30+
                 Windgeschwindigkeit100MP40+
                 Windgeschwindigkeit100MP50+
                 Windgeschwindigkeit100MP70+
                 Windgeschwindigkeit100MP90+
                 Verfügbare_Kapazität +
                 month + 
                 winddirectionvar + 
                 windspeedvar +
                 winddirectionweightedmean,
               mtry = 8,
               data=train)
plot(rforest.2)
```

#### Predictions
```{r Preds_Trf}
predict.rforest2 <- round(predict(rforest.2, newdata=eval), digits=0)
(sum((eval$Output-predict.rforest2)^2)/nrow(eval))^(1/2)
error_trf <- (sum(abs(predict.rforest2-eval$Output)))/sum(eval$Verfügbare_Kapazität)
error_trf

results <- 
  rbind(results, 
        list("Tuned Random Forest", 
             error_trf))
```

#### PLots
```{r Plots}
plot(rforest.2, col = "red", main = "Random Forest Final Model Comparison")
plot(rforest.1, col= "blue", add= TRUE)
plot(rforest, col = "green", add= TRUE)
```

### Saving Results
```{r Save_Results}
save(results, "Data_out/results.Rda")

```